---
title: "CrimeRateVsRepairTime"
author: "Virginia Saulnier"
date: "June 20, 2016"
output: html_document
---
This program takes in data for Chicago crime rates, alley lights out, and population per district and compares these factors.


Create a function to read in files, also has ability to skip extra header line, uses headers as column titles and returns data
```{r}

library(dplyr)
library(Runit)

readin <- function(wd,file,skip){
  setwd(wd)
  dataframe = readLines(file)
  if (skip == "TRUE")
  {
    dataframe = dataframe[-1]
  }
  
  data = read.csv(textConnection(dataframe),header=T, fill = TRUE )
  
  return(data)
}
```

Read in 3 data sets using function "readin".

```{r}

alleylights <- readin('/Users/VirginiaSaulnier/Desktop/412_hw1','311_Service_Requests_-_Alley_Lights_Out.csv',TRUE)
crime <- readin('/Users/VirginiaSaulnier/Desktop/412_hw1','Crimes2016.csv',FALSE)
population <- readin('/Users/VirginiaSaulnier/Desktop/412_hw1','District_Population.csv',FALSE)

```

Assigns column names for dataframe "alleylights",filters only repairs from 2016 and creates new dataframe "alley16", caculates repair time by and puts repair time in new column "date_diff"

```{r}
colnames(alleylights) <- c("Request","Status","Completion", "Number","Type", "Address", "Zip", "1","2","Ward","District","Community", "Latitude","Longitude","Location")

alley16 <- alleylights[grep("2016", alleylights$Request), ]
alley16$date_diff <- as.Date(as.character(alley16$Completion),format="%m/%d/%y")-
  as.Date(as.character(alley16$Request), format="%m/%d/%y")

# test.date_diff<- function(){
# checkTrue(is.date(date_diff))
# }
```

Creates 2 new dataframes seperating data by district and calculates the average repair time and crime counts by district.

```{r}
service_ave = alley16 %>%
  group_by(District) %>%
  summarise(repairtime = round(mean(date_diff, na.rm=TRUE),digits = 1)) %>%
  arrange(District) 
group_by(alley16, District)

crime_counts = crime %>% 
  group_by(District) %>%
  summarise(crimetotal=length(Primary.Type)) %>%
arrange(District)

```

Joins all 3 data sets

```{r}
results = full_join(crime_counts, service_ave, by = "District")%>%
  arrange(crimetotal)
results = full_join(results, population, by = "District")%>% 
  arrange(crimetotal)

```

Calculates a crime rate by dividing number of crimes per district by the population in each district. Plots results

```{r}
results$crimerate <- results$crimetotal/results$Population * 100
plot(results$repairtime, results$crimerate)

fit <- lm(results$repairtime ~ results$crimerate)
#abline(lm(repairtime~crimerate))

```



